---
title: "pinheiro"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{pinheiro}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
# packages used in this vignette
library(varameta) # variance estimator for the sample median
library(metafor) # for meta-analysing
library(kableExtra) # for pretty tables
library(tidyverse) # datasci toolkit
library(metaviz) # suggested by W. Kyle Hamilton
# library(simeta) # must move the dataset
conflicted::conflict_prefer("filter", "dplyr")

```

`varameta::` contains an example [meta-analysis dataset](https://www.sciencedirect.com/science/article/pii/S0009898112004007) [citation] that contains studies that report both means and medians.

```{r}
pinheiro_dat %>% 
  kable() %>% 
  kable_styling()

```

# calculate the standard error of means and medians

```{r}

rma_dat <-
  pinheiro_dat %>%
  mutate(
    logratio_effect = log(m_t / m_c),
    m_t_se = pmap_dbl(
      list(
        centre = m_t,
        spread = s_t_d,
        n = n_t,
        centre_type = centre,
        spread_type = spread
      ),
      effect_se
    ),
    m_c_se = pmap_dbl(
      list(
        centre = m_c,
        spread = s_c_d,
        n = n_c,
        centre_type = centre,
        spread_type = spread
      ),
      effect_se
    ),
    logratio_se = if_else(
      centre == "median",
      # variance of log-ratio of medians
      m_t_se/m_t^2 + m_c_se/m_c^2,
      # variance of log-ratio of means
      1/n_t*(s_t_d/m_t)^2 + 1/n_c*(s_c_d/m_c)^2
    ),
    study_label = str_c(study, year, sep = " ")
  ) 

# now we have the log-ratio of effects
# and estimates of the variance of the log-ratio
rma_dat %>% 
  select(study_label, year, logratio_effect, logratio_se)



```

Now that we have calculated the desired effect, $\log(M_1/M_2)$ where $M$ might denote mean or median for a given study, and the numbers denote the arms of the study, we can meta-analyse with `metafor::`.

# mean difference


```{r}
png("pinheiro-means.png")

pinheiro_dat %>%
  filter(centre == "mean") %>%
  mutate(study_label = str_c(study, year, sep = " ")) %>% 
  escalc(
    measure = "MD",
    m1i = m_t,
    n1i  = n_t,
    sd1i = s_t_d,
    m2i = m_c,
    n2i  = n_c,
    sd2i = s_c_d,
    slab = study_label,
    data = .
  ) %>% 
  rma(yi = yi, vi = vi, data = .) %>% {
    forest(., refline = .$beta)
  }

dev.off()
```

# median difference


```{r}

median_md_dat <- 
  pinheiro_dat %>%
  filter(centre == "median") %>%
  mutate(m_t_se = pmap_dbl(
    list(
      centre = m_t,
      spread = s_t_d,
      n = n_t,
      centre_type = centre,
      spread_type = spread
    ),
    effect_se
  ),
  m_c_se = pmap_dbl(
    list(
      centre = m_c,
      spread = s_c_d,
      n = n_c,
      centre_type = centre,
      spread_type = spread
    ),
    effect_se
  ),
  study_label = str_c(study, year, sep = " "),
         # calculate mean difference
         md = m_t - m_c,
         # calculate variance of mean difference

           md_v = m_t_se^2 + m_c_se^2)  

png("pinheiro-medians.png")

rma(yi = md, sei = sqrt(md_v), data = median_md_dat, slab = study_label) %>% {
  forest(., refline = .$beta)  
}

dev.off()

```

# experimenting

```{r eval=FALSE}
pinheiro_rma_means <- 
  rma(
    yi = logratio_effect,
    sei = logratio_se,
    data = rma_dat %>% filter(centre == "mean"),
    test = "knha",
    slab = study
  )

pinheiro_rma_medians <- 
  rma(
    yi = logratio_effect,
    sei = logratio_se,
    data = rma_dat %>% filter(centre == "median"),
    test = "knha",
    slab = study
  )



```


```{r eval=FALSE}

# copying this code wholesale from 
# https://rpubs.com/mcmurdie/ggforest
# suggested by Matt Grainger
ggforest = function(x, centre = "medians"){
  require("ggplot2")
  # Function to convert REM results in `rma`-format into a data.frame
  rma2df = function(x){
    rbind(
      data.frame(Study = "RE Model", LogFC = x$b, CILB=x$ci.lb, CIUB=x$ci.ub,
                 p = x$pval,
                 stringsAsFactors = FALSE),
      data.frame(Study = x$slab, LogFC = x$yi, 
                 CILB=x$yi - 2*sqrt(x$vi),
                 CIUB=x$yi + 2*sqrt(x$vi), 
                 p = x$pval,
                 stringsAsFactors = FALSE)
    )
  }
  remresdf = rma2df(x)
  remresdf <- transform(remresdf, interval = CIUB - CILB)
  remresdf <- transform(remresdf, RelConf = 1/interval)
  p = ggplot(remresdf, 
             aes(LogFC, Study, xmax=CIUB, xmin=CILB)) + 
    #coord_cartesian(xlim=c(-2, 2)) +
    #scale_alpha_discrete(range = c(0.2, 1)) +
    geom_vline(xintercept = 0.0, linetype=2, alpha=0.75) +
    geom_errorbarh(alpha=0.5, color="black") + 
    geom_point(aes(size = RelConf)) +
    geom_point(data = subset(remresdf, Study=="RE Model"), size=7) +
    scale_size(range = c(2, 5), guide=FALSE) +
    theme_bw() + 
    theme(text = element_text(size=16)) +
    labs(x = sprintf("log-ratio of %s", centre))
  return(p)
}
```



```{r fig.width=6, eval=FALSE}

# Matt's suggestion
pinheiro_rma_means %>% 
  ggforest(centre = "means", transf = exp)

pinheiro_rma_medians %>% 
  ggforest(centre = "medians")


# Kyle's suggestion
pinheiro_rma_means %>% 
  viz_forest()

pinheiro_rma_medians %>% 
  viz_forest()

# from November 2017!
pinheiro_rma_means %>% 
  prettyforest::ggforest()

pinheiro_rma_medians %>% 
  prettyforest::ggforest()


```

```{r eval=FALSE}

forest(pinheiro_rma_means, 
       refline = ex, 
       transf = exp, 
       digits = 4)




```



```{r}



```




