---
title: "varameta"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{varmeta}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
set.seed(39)
library(varameta)
```

## objective

The `varameta::` package aims to bridge the toolchain gap from a dataset containing medians to a software package, such as `metafor::`, for meta-analysis.

## calculate standard error of mean or median based on effect, spread, and sample size

A wrapper function `effect_se` provides a quick method of calculating the error of an effect based on its measure of effect, spread, sample size. 

```{r mean}
# consider a randomly generated sample
(a_sample <- rlnorm(10, 4, 0.2))

# summary statistics for this sample
a_sample %>% log() %>% summary()

# calculate the standard error based on the median, interquartile range, and 
# sample size

effect_se(
  centre = median(a_sample),
  spread = IQR(a_sample),
  n = length(a_sample),
  centre_type = "median",
  spread_type = "iqr"
)

```


### standard error of the sample median

### standard error of the sample mean

- explicit example
- then with sampling


## variance of the sample median

- explicit example
- then with sampling


## vectorised calculations

## other estimators

All other estimators available for meta-analysis are provided in `varameta::` for use in a compative analysis. 
